---
output:
  pdf_document: default
  html_document: default
---
# Statistical_Inference Course Project Part 2: ToothGrowth Dataset Analysis
## Author: Michael Nichols

Basic Summary of the Dataset:

ToothGrowth - data frame with 60 observations on 3 variables
- len	numeric	Tooth length
- supp	factor	Supplement type (VC or OJ).
---VC: ascorbic acid tablet
---OJ: orange juice
- dose	numeric	Dose in milligrams/day
```{r BasicSummary}
#load libraries
library(ggplot2)
library(datasets)
data("ToothGrowth")

str(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)

#basic plot of the summary statistics
qplot(factor(dose), len, data = ToothGrowth, color = factor(dose), facets = supp ~ ., geom = c("boxplot", "jitter"))
```

2. T-Test Evaluation

```{r ConfInf_HypothesisTests}
# Explore effects of supplement type on tooth growth by performing a two-sample t-test for the difference in tooth length by supplement (without including the dosage). We assume a confidence level of 95% and unequal variances.

ToothTTest <- t.test(len ~ supp, data = ToothGrowth, var.equal = FALSE, paired = FALSE)
```

3. Conclusions:

```{r Conclusions}
#show table of two sample t-test results for tooth growth by supplement
tTestResults <- matrix(data.frame(ToothTTest$p.value, ToothTTest$conf.int[1], ToothTTest$conf.int[2], ToothTTest$estimate[1], ToothTTest$estimate[2]))
rownames(tTestResults) <- c("p-value", "Lower Conf Int", "Upper Conf Int", "OJ Mean", "VC Mean")
colnames(tTestResults) <- "Results"

print(tTestResults)
```
We do not reject the Null hypothesis, as the p.value is 0.061 which is greater than the 0.05 threshold and the confidence interval contains 0. 